Traversing symmetric NAT with predictable port allocation

### 2014-07-15 NAT_paper_ACM_short.tex
ACM version for SIN2014

% \documentclass[twoside]{article}
\documentclass{llncs}
%\usepackage{llncsdoc}

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}

% ------
% Page layout
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
\usepackage{multicol}

@@

% ------
% Abstract
\usepackage{abstract}
	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

% ------
% Titling (section/subsection)
% \usepackage{titlesec}
% \renewcommand\thesection{\Roman{section}}
% \titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{}

% ------
% Header/footer
% \usepackage{fancyhdr}
% 	\pagestyle{fancy}
% 	\fancyhead{}
% 	\fancyfoot{}
% 	\fancyhead[C]{SPW 2014 $\bullet$ Twenty-second International Workshop on Security Protocols}
% 	\fancyfoot[RO,LE]{\thepage}
% CFP: http://www.wikicfp.com/cfp/servlet/event.showcfp?eventid=33620&copyownerid=44118

@@

% ------
% Math
% \usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
\usepackage{wrapfig}
\usepackage{tikz}
\usepackage{float}
\usepackage{multirow}
%\usepackage{appendices}
\usepackage{subcaption}

@@

% ------
% Maketitle metadata
\title{\vspace{-15mm}%
	\fontsize{24pt}{10pt}\selectfont
	\textbf{Traversing symmetric NAT with predictable port allocation function}
	}	
% \author{%
% 	\large
% 	\textsc{Du\v{s}an Klinec} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
% 	\normalsize	Faculty of Informatics \\
% 	\normalsize	Masaryk University, Brno \\
% 	\normalsize	\href{mailto:xklinec@fi.muni.cz}{xklinec@fi.muni.cz}
% 	\and
% 	\textsc{Vashek Matyáš} \\[2mm]%\thanks{Template by \href{http://www.howtotex.com}{howtoTeX.com}} \\[2mm]
% 	\normalsize	Faculty of Informatics \\
% 	\normalsize	Masaryk University, Brno \\
% 	\normalsize	\href{mailto:matyas@fi.muni.cz}{matyas@fi.muni.cz}
% 	\vspace{-5mm}
% 	}
% \date{}

\urldef{\mailsa}\path|xklinec@fi.muni.cz|
\urldef{\mailsb}\path|matyas@fi.muni.cz|
\author{Du\v{s}an Klinec \and Vashek Matyáš}
\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic\\
\email{\{xklinec,matyas\}@fi.muni.cz}}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\maketitle
% \thispagestyle{fancy}

\begin{abstract}
\noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
Presence of such (potentially malicious) hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
and security consequences.
Several solutions exist for traversing NAT, but symmetric case is still problematic. We propose algorithms using single source port for 
symmetric NAT traversal, each with different properties and applicability. 
Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
NAT as a random process, in order to evaluate algorithms in different settings.
\end{abstract}

% \begin{multicols}{2}
% \lettrine[nindent=0em,lines=3]{L}orem ipsum dolor sit amet.

\section{Introduction}
Network Address Translators (abbreviated as NATs) have been increasing in popularity in recent years, 
mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
the ability to share a public, globally routable IP address among many internal network hosts. On 
the other hand, it breaks the basic end-to-end principle of the Internet and thus favors server-client 
communication model for clients connected behind NAT.

In many applications, a peer-to-peer (P2P) connection is of crucial importance and the quality of a service depends on 
its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible, relay 
servers have to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on the communication channel quality 
(increased latency, jitter, lower bandwidth), the relay server is a potential point of failure and 
has to scale with increasing demands of the network, etc.
Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service providers where establishing a direct connection between parties is of high importance.

From the attacker's perspective, relay servers can be seen as potential allies when trying to sniff and log all 
communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
forces communicating parties to use such central servers
for communication, making monitoring easier for an attacker. Traversal algorithms have to collaborate 
with central servers in such a way they create a direct link, not using servers any further. 
% 
% some way. They use it to create a direct link, not using servers any further.

In this paper, we focused on a symmetric NAT with a predictable port allocation function. We propose 
new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
have low system requirements.

\section{Existing techniques}
Many different solutions for traversing non-symmetric NATs exist e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
Interactive Connectivity Establishment~(ICE)~\citep{rfc5245}. %Terredo\ignore{tunneling}~\citep{rfc4380}.

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. Yet in general, it is not widely deployed and it cannot be relied upon in a general case.

ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
although it does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.

% \section{Related work}
Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, while the case where both clients are behind symmetric NAT
with a port sensitive allocation function is not covered. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
Our motivation is to show that it is actually possible.

RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
The research by Wei~\emph{et~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared to ours, and the the downside of their algorithm comes with the Time To Live~(TTL) parameter modification of the IP~packet. This is usually considered
as a privileged operation in the operating system and super-user privileges are required for this modification.

Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
behind symmetric NATs, based on a changing of a source port, while we explore options with a constant source port.

In our work we study the situation where two parties -- $A, B$ -- try to establish a direct connection, both of them are behind 
a symmetric NAT (possibly a cascade of such NATs) with a port sensitive, predictable allocation function. Other scenarios with
predictable allocation function are quite easy to solve and covered in previous research.

@@

\begin{mydef}
% \begin{align*}
$\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket. $\mathbb{IP}$ is the set of all possible IP addresses and $\mathbb{PORT}$ 
is the set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
% \begin{align*}
$\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket Pair.
\end{mydef}
% 
% \begin{mydef}
% \begin{align*}
% \mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
% \end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
% \end{mydef}
% Allocation table is a set of all socket pairs and corresponding allocated external port.

\paragraph{NAT and communication.}
NAT connects two networks, denoted as {\em internal} and {\em external}, connected to an internal and an external NAT interface, respectively.
Hosts have addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network,
respectively, where $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
If it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
$\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.

@@

\subsection{NAT classes}\label{sec:nat}
NAT is classified w.r.t. the \emph{filtering rule} and the \emph{allocation rule}, which state what happens after receiving
packet from an external and an internal network, respectively.

@@

\paragraph{Port Restricted Cone NAT} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between the internal host socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
The filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Symmetric NAT} ~\\
Allocation table $\mathbb{AT}$ and filtering rule are the same as in the Port Restricted Cone NAT. The difference lies in the way 
the new entries in the allocation table are created. For the Cone NAT holds that multiple entries in the allocation table 
can have the same external port assigned, while the Symmetric NAT creates new allocation entries under certain circumstances.

\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be either \emph{predictable} or \emph{random}. In our setting, a predictable AF is assumed, e.g., \emph{incremental}. 
If the previously allocated port was $\text{PORT\textsubscript{nat}}$ then
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where typically $\Delta=1$.
If the chosen port is already taken, AF iterates until a free one is found.

AF can have different \emph{sensitivity} determining the cases where new allocation is created or the previous one is used instead.

\emph{Address sensitive allocation function} creates new allocation if a destination address differs from the existing allocation 
from the same source:
\begin{align*}
& \exists s=(a_1,a_2,a_3,a_4) \in \mathbb{SP}, p \in \mathbb{PORT}: \\
& (s, p) \in \mathbb{AT} \Rightarrow ((\forall s^{\prime}=(a_1^{\prime},a_2^{\prime},a_3^{\prime},a_4^{\prime}) \in \mathbb{SP}, \\
& a_1 \neq a_1^{\prime}, a_2 \neq a_2^{\prime}, a_3 \neq a_3^{\prime}, \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}

\emph{Port sensitive allocation function} creates new allocation if a destination address and a port differs from the existing
allocations from the same source:
\begin{align*}
& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow \\
& \Rightarrow ((\forall s^{\prime} \in \mathbb{SP}, s^{\prime} \neq s , \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}
% 
%     PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
% \end{align*}

% \section{Related work II}
% \begin{compactitem}
% \item chineese\citep{Wang:2006:RSN:1156422.1156550}
% \item uPNP
% \item PCP~\citep{rfc6887}
% \end{compactitem}

\section{UDP hole punching with STUN}
All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
Internet hosts cannot send packets to the internal host directly. 
Communication has to be initiated from the inside\footnote{Not taking
pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, which is problematic if two nodes behind 
NAT want to communicate.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B}, respectively.
A and B are behind a public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$,
respectively. By default an internal host does not know its external address and NAT type, thus a technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if Il get it, etc.} 
is used to learn it. In case of incremental NAT it is also necessary to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.

@@

\section{Theoretical network model}
In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
we decided to model network behavior with a stochastic process. In particular, we chose a Poisson process, $\{N(t), t\geq0\}$ 
that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.

The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
it is important to model how the NAT internal state changes over time, i.e., how many external ports 
have been allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.

We model the internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
different workloads of the network and evaluate algorithms w.r.t. various parameters, e.g., 
probability of establishing a connection.

@@

\begin{mydef}
$N(t)$ is a random variable of a number of a new port allocations made in time 
interval~$[0,t]$ where $N(t) \sim Po(\lambda t)$. \\

\begin{center}                                                     
Then $P[N(t)=n] = \frac{(\lambda t)^n}{n!} e^{-\lambda t}$
\end{center}
\end{mydef}

Moreover, we assume an incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
state\footnote{STUN requests to determine current port number, it creates a new NAT allocation record.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
which is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
If it is higher then NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.

\section{Generic algorithm structure}
Algorithms that we propose have a common structure. In order to maximize the probability of establishing
a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
tries to open a new hole in the local NAT and/or to predict the corresponding port allocated in the destination
NAT.

>>>>

\documentclass{acm_proc_article-sp}
% http://www.acm.org/sigs/publications/proceedings-templates

% ------
% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% \linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}

% ------
% Page layout
% \usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage[font=it]{caption}
\usepackage{paralist}
% \usepackage{multicol}

@@

% ------
% Abstract
% \usepackage{abstract}
% 	\renewcommand{\abstractnamefont}{\normalfont\bfseries}
% 	\renewcommand{\abstracttextfont}{\normalfont\small\itshape}

@@

% ------
% Math
% \usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
\usepackage{wrapfig}
% \usepackage{tikz}
\usepackage{float}
\usepackage{multirow}
%\usepackage{appendices}
\usepackage{subcaption}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{Traversing symmetric NAT with predictable port allocation}

\numberofauthors{1}
\author{
% 1st. author
\alignauthor
Du\v{s}an Klinec, Vashek Maty\'{a}\v{s}\\
       \affaddr{Faculty of Informatics, Masaryk University}\\
       \affaddr{Brno, Czech Republic}\\
       \email{\{xklinec,matyas\}@fi.muni.cz}
% % 2nd. author
% \alignauthor
% G.K.M. Tobin\titlenote{The secretary disavows
% any knowledge of this author's actions.}\\
%        \affaddr{Institute for Clarity in Documentation}\\
%        \affaddr{P.O. Box 1212}\\
%        \affaddr{Dublin, Ohio 43017-6221}\\
%        \email{webmaster@marysville-ohio.com}
% % 3rd. author
}

\maketitle
\begin{abstract}
\noindent Network Address Translators often cause trouble for VoIP and other P2P services since central servers are needed for communication.
The presence of such potentially malicious hosts in a communication path is not desired, mainly due to security consequences, poor link quality 
and increased cost.
Several solutions exist for traversing NAT, but a symmetric one is still problematic. We propose algorithms using a single source port for 
symmetric NAT traversal. Each with different properties and applicability. 
% Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
% NAT as a random process, in order to evaluate algorithms in different settings.
\end{abstract}

\section{Introduction}
Network Address Translators (NATs) have been increasing in popularity in recent years, 
mainly as a means of slowing down IPv4 address space depletion. On one hand, NAT offers
the ability to share a public, globally routable IP address among many internal network hosts. On 
the other hand, it breaks the basic end-to-end principle of the Internet and thus favors server-client 
communication model for clients connected behind NAT.

In many applications, a peer-to-peer (P2P) connection is of crucial importance and the quality of a service depends on 
its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible to establish, relay 
servers have to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on the communication channel quality 
such as increased latency, jitter, lower bandwidth. The relay server is a potential single point of failure and 
has to scale with increasing demands of the network, etc.
% Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service 
% providers where establishing a direct connection between parties is of high importance.

In this paper, we focused on a symmetric NAT with a predictable port allocation function. 
The contributions of this paper are as follows:%
% \begin{compactitem}
% \item 
a) we propose new effective algorithms for establishing a direct connection between hosts behind such NATs; %
% \item 
b) we demonstrate the implementations of these algorithms; %, together with a simulation based on a theoretical approach from a queueing theory. These implementations we provide in public domain.%
% \item 
c) we tested our algorithms using data from a real network; % (of our mobile phone operator).%
% \item 
d) last but not least, we provide a formalization of the NAT traversal problem.
% \end{compactitem}
% We propose 
% new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
% have low system requirements.

\section{Existing techniques}
Many different solutions for traversing non-symmetric NATs exist, e.g., NAT traversal in Internet Key Exchange~\citep{rfc3947} or 
Interactive Connectivity Establishment~(ICE)~\citep{rfc5245}. %Terredo\ignore{tunneling}~\citep{rfc4380}.

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. However, in general, 
it is not widely deployed and it cannot be relied upon in a general case. 
Another solution is ICE\ignore{~\citep{rfc5245}}. It is used as a technique to traverse NAT in most applications nowadays 
although it does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.

% \section{Related work}
Interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, although does not cover the case with both clients using symmetric NAT
with a port sensitive allocation function. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
We show that it is actually possible. Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
behind symmetric NATs, based on changing of a source port, while we explore options with a constant source port.

RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
The research by Wei~\emph{et~al.} in \citep{wei} is devoted to symmetric NAT traversal, but their algorithm is using a different from ours.
Disadvantage of their approach comes with the Time To Live~(TTL) parameter modification of the IP~packet. This is usually considered
as a privileged operation in the operating system and super-user privileges are required for this modification.

In our work we study the situation where two parties, $A$ and $B$, try to establish a direct connection, both of them are behind 
a symmetric NAT (possibly a cascade) with a port sensitive, predictable allocation function. 
%Other scenarios with predictable allocation function are quite easy to solve and covered in previous research.

@@

\begin{mydef}
% \begin{align*}
$\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket. $\mathbb{IP}$ is the set of all possible IP addresses and $\mathbb{PORT}$ 
is the set of all possible ports. %, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
% \begin{align*}
$\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket Pair.
\end{mydef}
% 
% \begin{mydef}
% \begin{align*}
% \mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
% \end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
% \end{mydef}
% Allocation table is a set of all socket pairs and corresponding allocated external port.

% \paragraph{NAT and communication.}
\subsection{NAT and communication}
NAT connects two networks, denoted as {\em internal} and {\em external}, connected to an internal and an external NAT interface, respectively.
Hosts have addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network,
respectively, where $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
Unless it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
$\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.

@@

\subsection{NAT classes}\label{sec:nat}
NAT is classified with respect to a \emph{filtering rule} and a \emph{allocation rule}, which state what happens after receiving
packet from an external and an internal network, respectively.

@@

% \paragraph{Port Restricted Cone NAT} ~\\
\par\smallskip
\noindent\textbf{Port Restricted Cone NAT.} 
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between the internal host socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
The filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

% \paragraph{Symmetric NAT} ~\\
\par\smallskip
\noindent\textbf{Symmetric NAT.} 
Allocation table $\mathbb{AT}$ and filtering rule are the same as in the Port Restricted Cone NAT. The difference lies in the way 
the new entries in the allocation table are created. For the Cone NAT holds that multiple entries in the allocation table 
can have the same external port assigned, while the Symmetric NAT creates new allocation entries under certain circumstances.

\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be either \emph{predictable} or \emph{random}. In our setting, a predictable AF is assumed, e.g., \emph{incremental}. 
If the previously allocated port was $\text{PORT\textsubscript{nat}}$ then
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where typically $\Delta=1$.
If the chosen port is already taken, AF iterates until a free one is found.
AF can have different \emph{sensitivity} determining the cases where new allocation is created or the previous one is used instead.

\emph{Address sensitive allocation function} creates a~new allocation if a destination address differs from the existing allocation 
from the same source:
\begin{align*}
& \exists s=(a_1,a_2,a_3,a_4) \in \mathbb{SP}, p \in \mathbb{PORT}: \\
& (s, p) \in \mathbb{AT} \Rightarrow ((\forall s^{\prime}=(a_1^{\prime},a_2^{\prime},a_3^{\prime},a_4^{\prime}) \in \mathbb{SP}, \\
& a_1 \neq a_1^{\prime}, a_2 \neq a_2^{\prime}, a_3 \neq a_3^{\prime}, \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}

\emph{Port sensitive allocation function} creates a new allocation if a destination address and a port differs from the existing
allocations from the same source:
\begin{align*}
& \exists s \in \mathbb{SP}, p \in \mathbb{PORT}: (s, p) \in \mathbb{AT} \Rightarrow \\
& \Rightarrow ((\forall s^{\prime} \in \mathbb{SP}, s^{\prime} \neq s , \forall p^{\prime} \in \mathbb{PORT}: \\
& ((s^{\prime}, p^{\prime}) \in \mathbb{AT})) \Rightarrow p \neq p^{\prime})
\end{align*}
% 
%     PASS((\text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}} )) \Leftrightarrow \\
% &  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
% & ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{out}}, \text{PORT\textsubscript{out}}), \text{PORT\textsubscript{ex}} ) \in \mathbb{AT}
% \end{align*}
% \section{Related work II}
% \begin{compactitem}
% \item chineese\citep{Wang:2006:RSN:1156422.1156550}
% \item uPNP
% \item PCP~\citep{rfc6887}
% \end{compactitem}

\section{UDP hole punching with STUN}
All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
A public internet host cannot send packets to the private host behind NAT/firewall directly. 
Communication has to be initiated from the inside\footnote{Not taking
pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, what is problematic if both
communicating sides are private.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B}, respectively.
A and B are behind a public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$,
respectively. By default an internal host does not know its external address or NAT type, thus a technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if Il get it, etc.} 
is used to learn it. In case of incremental NAT it is also necessary to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending it to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.

@@

\section{Theoretical network model}
In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
we decided to model network behavior with a stochastic process. In particular, we chose a \emph{Poisson process}, $\{N(t), t\geq0\}$ 
that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.

The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
it is important to model how the NAT internal state changes over time, i.e., how many external ports 
were allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.

We model the internal network connected to NAT with respect to newly created connections (a new port allocation) as a 
\emph{time homogeneous Poisson process} $Po(\lambda)$. Value of $\lambda$ models
different workloads of the network. This way we can evaluate algorithms with respect to various parameters, e.g., 
probability of establishing a connection.

@@

\begin{mydef}
$N(t)$ is a random variable of a number of new port allocations made in time 
interval~$[0,t]$ where $N(t) \sim Po(\lambda t)$.
\begin{center}                                                     
Then $P[N(t)=n] = \frac{(\lambda t)^n}{n!} e^{-\lambda t}$
\end{center}
\end{mydef}

Moreover, we assume an incremental NAT with $\Delta=1$ (without loss of generality) on both sides. If we sample the NAT 
state\footnote{STUN requests to determine current port number, it creates a new NAT allocation record.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

% \paragraph{Port pool exhaustion.} 
\par\smallskip
\noindent\textbf{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
what is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~sec on average. 
If it is higher then NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.

\section{Generic algorithm structure}
Algorithms that we propose have a common structure. In order to maximize the probability of establishing
a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
tries to open a new hole in the local NAT and to predict the corresponding port allocated in the destination
NAT.

\begin{mydef}
The symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where 
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is the set of communicating parties.
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
\item $\mathcal{M}$ is the set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) $ $| \lambda~>~0, t\geq~0 \}$. Value $t_s$ denotes the length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively.
\end{compactitem}
\end{mydef}
If $\mathcal{M}$ is not important for the algorithm it can be ommitted in the notation.
%If the third parameter is not important for the algorithm it can be ignored in the notation.

Prior algorithm run it is needed to 
determine the NAT state (for UDP hole punching), this is assumed to be done by a technique like STUN. For simplicity, we assume that port numbers
start from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
the NAT, e.g., difference between the last measurement and just before actual algorithm execution.

% \paragraph{Notation.}
\par\smallskip
\noindent\textbf{Notation.} 
For instance, line $A(0,1) = (2,3)$ means that party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party~$1$.
